Temporal stability in human interaction networks

### 2015-08-13 paper.tex
adicionadas modificacoes entregues pelo leo

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution was investigated e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Notably, the activity in terms of seconds, minutes, hours, days and months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing against the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{newmanBook}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.

@@

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a unit complex number, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{period}$. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
\end{align}

@@

Edges can be created from all antecedent message authors on the message-response thread to each message author.
We only linked the immediate antecedent to the new message author, both for simplicity and because in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$. Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''. This suggests the adequacy of centrality measurements to account for the connectivity of a node with all other nodes, such as betweenness centrality, eigenvector centrality and accessibility~\cite{luMeasures,access}.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\textwidth]{figs/fser_}
	\caption{Classification of the scale-free network vertices by comparing the degree
		distribution against that of an Erd\"os-R\'enyi ideal network.
		The latter has more intermediary vertices, while the former has more peripheral and hub vertices.
		The sector borders are defined by the two intersections $k_L$ and $k_R$ of the connectivity distributions.
		Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
		\label{fig:setores}
\end{figure}
To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network. Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j})$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than $k_{j}$. This changes equation~\ref{criterio} to

\begin{equation}\label{criterio2}
	\sum_{x=k_i}^{k_j} \widetilde{P}(x) < \sum_{x=k_i}^{k_j} P(x) \Rightarrow \text{i is intermediary.}
\end{equation}

>>>>

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution was investigated e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Notably, the activity in timescales ranging from seconds to months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing against the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space are centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{moreno}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.

@@

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a unit complex number, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{T}$, where T is the period in which the counting is repeated. E.g. $\theta=12\frac{2\pi}{24}=\pi$ for a message sent at $12h$ and given $T=24h$ for days. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{T}{2\pi} \theta_\mu \nonumber
\end{align}

@@

%Edges can be created from all antecedent message authors on the message-response thread to each message author.
%We only linked the immediate antecedent to the new message author, both for simplicity and because in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$. Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''. This suggests the adequacy of centrality measurements to account for the connectivity of a node with all other nodes, such as betweenness centrality, eigenvector centrality and accessibility~\cite{luMeasures,access}.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\textwidth]{figs/fser_}
	\caption{Classification of the scale-free network vertices by comparing the degree
		distribution against that of an Erd\"os-R\'enyi ideal network.
		The latter has more intermediary vertices, while the former has more peripheral and hub vertices.
		The sector borders are defined by the two intersections $k_L$ and $k_R$ of the connectivity distributions.
		Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
		\label{fig:setores}
\end{figure}
To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network. The range $\Delta$ of incident values should be partitioned in $m$ parts $\Delta=\cup_{i=1}^m \Delta_i$, with $\Delta_i\cap \Delta_j \; \forall\; i \neq j$. Thus, $\Delta_i=\{max(\Delta_{i-1})<k\leq j \;\;|\; \sum_{\Delta_{i-1}+1}^j \eta_k > \eta \}$, with $\eta_k$ the number of vertices with degree $k$ and $max(\Delta_{0})=-1$.This changes equation~\ref{criterio} to

\begin{equation}\label{criterio2}
	\begin{split}
	\sum_{x=min(\Delta_i)}^{max(\Delta_i)} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{max(\Delta_i)} P(x) \Rightarrow \\
	\Rightarrow \Delta_i \text{ hods intermediary degree values.}
	\end{split}
\end{equation}
